# 2016 Presidential Campaign Contributions in Michigan 

### by Liz Kalina
========================================================

## Preparations

#### Load Packages

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(gender)
library(tidyr)
library(devtools)
library(ade4)
library(data.table)
library(grid)
library(choroplethrZip)
```

#### Load Datasets

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Contribution Data and Column Names for Michigan
header <- read.table("2016-MI.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
contrib <- read.table("2016-MI.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(contrib) <- header

# Load Candidate Info (i.e. Party Affiliation)
header <- read.table("candidate-list.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
cand_info <- read.table("candidate-list.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(cand_info) <- header

# Load Population Density and Unemployment Information
header <- read.table("unemployment-zipcode.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
unemployment <- read.table("unemployment-zipcode.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(unemployment) <- header

header <- read.table("population-density-zipcode.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
pop_density <- read.table("population-density-zipcode.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(pop_density) <- header

# Remove NA column(s)
contrib <- contrib[colSums(!is.na(contrib)) > 0]
```

This report explores a dataset containing campaign contributions to all
candidates in the 2016 presidential election by Michigan residents. As a former 
Michigander, I'm particularly interested in Michigan's current political climate.

## Univariate Plots Section

#### Dataset Characteristics
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
str(contrib)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
length(unique(contrib$cand_nm))
```

Our dataset consists of 18 variables, with 163,765 observations and 24 
candidates that people contributed to.

#### Summary of Contributions to Each Candidates
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

cands <- group_by(contrib, cand_nm)
cand_summary <- summarize(cands, total = length(contb_receipt_amt),
                          avg = mean(contb_receipt_amt, na.rm = TRUE), 
                          median = median(contb_receipt_amt), 
                          max = max(contb_receipt_amt))

ordered_cands <- cand_summary[order(cand_summary$total, decreasing=TRUE) , ]
ordered_cands
```

The top Democratic candidates (Hillary Clinton, Bernie Sanders) received the 
largest number of donations, while the top three Republican candidates (Donald 
Trump, Ted Cruz, Ben Carson) received overall fewer donations but their median 
donations were roughly double that of the Democrats.

#### Summary of Contributions by City
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
city <- group_by(contrib, contbr_city)
city_summary <- summarize(city, total = length(contb_receipt_amt),
                          avg = mean(contb_receipt_amt, na.rm = TRUE), 
                          median = median(contb_receipt_amt), 
                          max = max(contb_receipt_amt))

city_summary[order(city_summary$total, decreasing=TRUE) , ]
```
The five cities with the largest number of donations were Ann Arbor, Grand 
Rapids, Detroit, Kalamazoo and Lansing.

### Contribution Histograms

#### Overall Distribution
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = contb_receipt_amt), data = contrib) + 
  geom_histogram(binwidth=5) +
  scale_x_sqrt(limits = c(0,2800))
```
A vast majority of donations were less than $500 which shows that at least in
the context of individual donations, it's not just rich donors making a majority
of the contributions.

#### Contributions less than $100
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt < 100)) + 
  geom_histogram(binwidth=10) +
  scale_x_continuous(limits = c(0, 100))
```
For donations less than $100, you can notice peaks around $25, $50 and $75
which makes sense since suggested donation amounts are often in $25 increments.

#### Contributions less than $400
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt < 400)) + 
  geom_histogram(binwidth=50) +
  scale_x_continuous(limits = c(0, 400))
```

#### Contributions greater than $1000
```{r echo=FALSE,message=FALSE, warning=FALSE,  Univariate_Plots}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt > 1000)) + 
  geom_histogram(binwidth=750) +
  scale_x_continuous(limits = c(1000, 6000)) +
  scale_y_continuous(limits = c(0, 1500))
```

According to the Federal Election Commission:
> $2,700 per election to a Federal candidate or the candidate's campaign committee. Notice that the limit applies separately to each election. Primaries, runoffs and general elections are considered separate elections.

Since presidential campaigns have primaries and general elections, you can see 
this reflected in the clustering of contribution amount at $2,700 and $5,400. 

### Candidates with the Most and Fewest Donations
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
p1 <- ggplot(aes(y=total, x= cand_nm),data=head(ordered_cands, 5)) + 
        geom_bar(stat="identity") + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle("Most Contributions")

p2 <- ggplot(aes(y=total, x= cand_nm),data=tail(ordered_cands, 5)) + 
        geom_bar(stat="identity") + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle("Least Contributions")


grid.arrange(p1,p2,ncol=2)
```
Above is a quick scale comparison of the candidates with the most and least
contributions, as you can see the scale difference is vast. Top candidates can
have more than half a million contributions while there are candidates on the
ballet who receive less than 10 inidvidual donations.


### Contributor Occupations and Employers

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
contrib %>% 
  group_by(contbr_employer) %>% 
  summarise(total = n()) %>%
  arrange(desc(total))
```
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
contrib %>% 
  group_by(contbr_occupation) %>% 
  summarise(total = n()) %>%
  arrange(desc(total))
```

Employment Observations
+ Retired people are the most freqeunt contibutors by far
+ Employees of universities, automakers and hosptial systems are frequent contributors
+ There are a large number of contributions without employer and occupation info


### Contribution Timing

#### Add Date to Contributions
```{r echo=FALSE, message=FALSE, warning=FALSE, Formatting}
contrib$contrib_date <- as.Date(contrib$contb_receipt_dt, '%d-%b-%y')
contrib$contrib_month_year <- format(contrib$contrib_date, '%Y-%m')
```

#### Contribution Totals by Date
```{r echo=FALSE, message=FALSE, warning=FALSE,  Univariate_Plots}

contrib_by_month_year <- contrib %>% 
  group_by(contrib_month_year) %>% 
  summarise(total = n())

ggplot(data=contrib_by_month_year, aes(x=contrib_month_year, y=total, group=1)) +
  geom_line() +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```
From the plot above, you can seek a gradual climb in donations during 2015, from
January 2016 to March 2016. There are three contribution peaks in 2016 in March,
July and October each coinciding with an election event; the Michigan Primary 
was March 8th, the Democratic and Republican national conventions occurred at the
end of July and the General Election was November 8th.

## Univariate Summary

To me, the most interesting features in this dataset are 'cand_nm', 'contbr_nm', 
'contbr_zip', 'contb_receipt_amt' and 'contb_receipt_dt.' I can chart 
'contb_receipt_amt' and 'contb_receipt_dt' for each contribution to investigate
temporal and candidate support trends. With 'contbr_nm' and 'contbr_zip', I can
infer information about the contributor such as geographic location and gender.

A few takeaways from my analysis so far are that Democratic candidates received
more contributions than Republicans and that a majority of contributions are
small (less than $500). Other interesting things I noticed were that contribution
amounts cluster around campaign maxs for larger donors and contributions peak 
around key election events (like primaries, conventions or scandals).

## Bivariate Plots Section

### Determine Contributor Gender and Candidate Political Party Affiliation

#### Add Political Party Affiliation
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
contrib <- merge(contrib, subset(cand_info[c("cand_id", "cand_pty_aff")]), by = "cand_id")
```
This information comes from the Federal Election Commission website.

#### Split Full Name column into First and Last Names
```{r echo=FALSE, message=FALSE, warning=FALSE, Formatting}
contrib <- separate(contrib, contbr_nm, c("last_name", "first_name"), sep = ",", extra="merge", fill="right")
contrib <- data.frame(lapply(contrib, trimws))
contrib$clean_first_name <- gsub("([A-Za-z]+).*", "\\1", contrib$first_name)
```

#### Add Gender to Contributors
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
gender_lookup <- gender(c(unique(contrib$clean_first_name)))
contrib <- left_join(contrib,gender_lookup,by = c("clean_first_name"="name"))

contrib_by_date_w_gender <- subset(contrib,proportion_male > 0.75 | proportion_female > 0.75) %>% 
  group_by(contrib_month_year, gender ) %>% 
  summarise(total = n())
```

In order to determine each contributors gender, I used the 'gender' package in R,
which encodes gender based on names and dates of birth using historical datasets.
I decided to use a threshold of 75% certainty to assign a gender to a name.

### Contributor Political Party Investigation

#### Contributor Gender vs. Political Party
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
ggplot(aes(x=cand_pty_aff),
       data=subset(contrib,!is.na(gender) & cand_pty_aff %in% c('Republican Party','Democratic Party'))) +
  geom_histogram(stat='count',aes(fill=cand_pty_aff)) +
  facet_wrap(~gender)
```
Both men and women in Michigan donate more to Democrats than Republicans,
however women are significantly more likely to donate Democratic than men.

#### Contributions Totals by Date and Political Party
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

contrib_by_date_w_party <- contrib %>% 
  group_by(contrib_month_year, cand_pty_aff ) %>% 
  summarise(total = n())

ggplot(data=na.omit(subset(contrib_by_date_w_party,cand_pty_aff != " ")), aes(x=contrib_month_year, y=total, group=cand_pty_aff)) +
  geom_line(aes(color = cand_pty_aff)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```
While Republicans actually donated more prior to 2016, the Democratic candidates
received the overall majority of contributions. There were very few donations
that were Libertarian or not marked for the a candidate.

#### Republican Candidate Contribution Totals by Date
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

reps <- subset(contrib, cand_pty_aff == 'Republican Party') %>% 
  group_by(contrib_month_year, cand_nm) %>% 
  summarise(total = n())

ggplot(data=subset(reps,total > 10), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```
I think this plot is really interesting because it shows that Trump had almost 
no financial contributions prior the Republican Primary on March 8, 2016. The 
leaders before the primary were Ben Carson and Ted Cruz. Trump's support 
peaked at the Republican National Convention at the end of July and then
steadily declined until the month before the general election when it had another
small resurgence.

## Bivariate Summary

In this section, I added political affiliation to candidates and gender to
contributors, which allowed me to see the relationship between these features
and campaign contributions. 

There is a very strong correlation between female contributors and Democratic 
contributions. Other discoveries include how cluttered the Republican 
race was, the strong Bernie support prior to the Democratic National Convention 
and the lack of Trump support prior to the Republican Primary.

## Multivariate Plots Section

#### Hillary vs. Trump over Time
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

hill_v_trump <- subset(contrib, cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.') & !is.na(gender)) %>% 
  group_by(contrib_month_year, gender, cand_nm) %>% 
  summarise(total = n())

ggplot(data=hill_v_trump, aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_bar(aes(fill = cand_nm),position = "dodge",stat="identity") +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~gender)

```
When looking at donations to Hillary vs. Trump, you can see that donation
trends are uniform across genders. However, women always donated more to 
Hillary than Trump, while men donated more to Trump in months leading up to the
election.


### Contribution Amount Distribution

#### Top 10 Candidates
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
top_10 <- head(ordered_cands, 10)$cand_nm

ggplot(aes(x = cand_nm,y=contb_receipt_amt), 
       data=subset(contrib,cand_nm %in% top_10 & contb_receipt_amt > 0 & contb_receipt_amt <= 2700)) + 
  geom_boxplot(aes(fill=cand_pty_aff)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Observations
+ The Republicans receive larger contributions than Democrats
+ Hillary has a low interquartile range but still have many large contributions 
+ Jeb Bush has comically large contributions 

#### By Political Party and Gender
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
contrib$contb_receipt_amt <- as.numeric(as.character(contrib$contb_receipt_amt))

ggplot(aes(x = cand_pty_aff,y=contb_receipt_amt), data=subset(contrib,!is.na(gender) & cand_pty_aff %in% c('Republican Party','Democratic Party') & contb_receipt_amt > 0 & contb_receipt_amt <= 2700)) + 
      geom_boxplot(aes(fill=cand_pty_aff)) +
      facet_wrap(~gender)
```
From this boxplot, you can see that the interquartile range for contribution 
amount is less than $100 across gender/political party and that Republican men 
give the most per contribution.

### Determine Population Density and Unemployment Rate of Contributor Zip Codes

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
contrib$contbr_zip <- as.integer(substr(contrib$contbr_zip, 0, 5))
contrib <- left_join(contrib,unemployment,by = c("contbr_zip"="Zip"))
contrib <- left_join(contrib,pop_density,by = c("contbr_zip"="Zip/ZCTA"))

#contrib$`Density Per Sq Mile` <- as.numeric(levels(contrib[,'Density Per Sq Mile']))[contrib[,'Density Per Sq Mile']]
contrib$`Unemp. Rate` <- as.numeric(sub("%", "", contrib$`Unemp. Rate`))
```
These two datasets were available on the Splitwise blog as downloads. According
to the author, this data came from:

> Two different Census APIs (the Decennial Census 2010 and the ACS 5-year 2007-2011), 
combined with the square-footage by ZCTA listings from the 2013 U.S. Gazetteer Files.

I believe that understanding the community type and economic prospects will be 
a strong indicator of candidate preference.

### Classify Contributor Community Type based on Population Density 

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
contrib$community_type <- cut(contrib$`Density Per Sq Mile`, breaks = c(0, 102, 2213,10000), labels = c('rural', 'suburban', 'urban'), right=FALSE)
```

Classification (Households per Square Mile):
+ rural: < 102
+ suburban: 103 to 2,213
+ urban: > 2213

These cutoffs are based on an analysis presented on the FiveThirtyEight blog:

> Our analysis showed that the single best predictor of whether someone said his or her area was urban, suburban or rural was ZIP code density. Residents of ZIP codes with more than 2,213 households per square mile typically described their area as urban. Residents of neighborhoods with 102 to 2,213 households per square mile typically called their area suburban. In ZIP codes with fewer than 102 households per square mile, residents typically said they lived in a rural area.


#### Party Affiliation vs. Unemployment vs. Population Density
```{r echo=FALSE,message=FALSE, warning=FALSE,Multivariate_Plots}
ggplot(aes(y=`Unemp. Rate`,x=`Density Per Sq Mile`), data=contrib) +
    geom_point(aes(color=cand_pty_aff),alpha = 1/100, size=3) +
      scale_color_brewer(type = 'div',
      guide = guide_legend(title = 'Party Affiliation', reverse = T,
      override.aes = list(alpha = 1, size = 2))) +
      scale_x_sqrt() +
      xlim(0,8000) +
      ylim(0,35)
```

In general, Democratic contributors appear to live in areas with higher population
density and higher unemployment while Republicans seem to live in areas with 
average unemployment and lower to moderate population density. What I would
takeaway from this is that Democratic contributors tend to live in urban areas
while Republican contributors inhabit more rural and suburban areas.

#### Correlation: Population Density, Gender, Party Affiliation
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

feats = c('gender','cand_pty_aff')
for (f in feats){
  df_dummy = acm.disjonctif(contrib[f])
  contrib = cbind(contrib, df_dummy)
}

cor_1 <- cor(na.omit(contrib[,c('Density Per Sq Mile','gender.male','gender.female','cand_pty_aff.Democratic Party','cand_pty_aff.Republican Party')]))

upper <- round(cor_1,2)
upper[upper.tri(cor_1)] <- ''
upper <- as.data.frame(upper)

upper
```
Observations
+ There are only minor correlations:
  + Democrats: higher population density, female
  + Republicans: lower population density, male


### Group by Party and Zip Code
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

## by zip code
group_by_zip <- contrib %>% 
  group_by(contbr_zip) %>% 
  summarise(total=n(),value=median(contb_receipt_amt))

setnames(group_by_zip, "contbr_zip", "region")
group_by_zip$region <- as.character(group_by_zip$region)

## by political party
group_by_party <- contrib %>%
 group_by(contbr_zip,cand_pty_aff) %>% 
 summarise(n= n()) %>%
 ungroup() %>%
 spread(cand_pty_aff, n, fill=0) %>%
 mutate(total = `Republican Party` + `Democratic Party`,
        rep_percent = (`Republican Party`/total),
        dem_percent = (`Democratic Party`/total)) %>%
 mutate(value = ifelse(`Republican Party`>`Democratic Party`, -(rep_percent),
               ifelse(`Republican Party`<`Democratic Party`, (dem_percent), NA)))

setnames(group_by_party, "contbr_zip", "region")
group_by_party$region <- as.character(group_by_party$region)

```

For the politcal party calculations, I select the dominant party (based on number
of contributions) for each zip code and calculate the percentage of total 
contributions that party receives.

#### Median Donation by Zip Code
```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

don_amt <- subset(group_by_zip,total > 10 & !is.na(region), c('region','value'))

zip_choropleth(don_amt, 
               state_zoom = "michigan", 
               title      = "Median Donation by Zip Code",
               legend     = "Median Donation",
               num_colors = 4) + coord_map()
```

The geography of median donation seems to be relatively evenly distributed across
the state with the exeception of Metro Detroit which is almost uniformly less 
than $30. The black regions are zip codes with less than 10 donations.

#### Correlation: Median Donation, Total Donations, Party Leaning

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

amt_pty_cor <- full_join(group_by_zip,group_by_party,by = 'region')

setnames(amt_pty_cor, "value.x", "median donation")
setnames(amt_pty_cor, "total.x", "total donations")

pairs(na.omit(amt_pty_cor[,c('median donation','total donations','dem_percent','rep_percent')]))
```

The above scatterplot matrix shows each zip code as a point. This plot simply
reinforces that Democrats live in more densely populated areas than Republicans
since the Democratic zip code have more total donations than the Republican zips.

## Multivariate Summary

In this section, I investigated support for Hillary vs. Trump over time, the
size of campaign contributions for Democrats vs. Republicans as well as the 
relationship between population density and unemployment to party affiliation.

I observed that Democrats indicators include high population density and female 
gender, while Republican indicators are low population density and male gender. 
Other observations include that Republicans (particularly Republican men) give 
larger donations and although men favored Trump to Hillary, men actually made 
more contributions to Hillary in the last few months leading up to the general 
election.

------

## Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

## Plot One

### Hillary vs. Bernie
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

hill_v_bern <- subset(contrib, cand_nm %in% c('Clinton, Hillary Rodham','Sanders, Bernard')) %>% 
  group_by(contrib_month_year, cand_nm) %>% 
  summarise(total = n())

ggplot(data=hill_v_bern, aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_bar(aes(fill = cand_nm),position = "dodge",stat="identity") +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)]) +
  ggtitle("Hillary vs. Bernie Support by Date")
  
```

### Description One
The above plot shows that Bernie Sanders actually more support (based on number
of contributions) from 2015 to well after the Democratic Primary. It was not
until June 2016 that Hillary surpassed Bernie in number of donations, and not
until after the Democratic convention that his support fully dried up. This
coincides with my understanding of Bernie as a grassroots candidate with
widespread and loyal supporters.

## Plot Two

### Party Leanings based on Zip Code

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

pty_ct <- subset(group_by_party,total > 5 & !is.na(region), c('region','value'))

zip_choropleth(pty_ct, 
               state_zoom = "michigan", 
               title      = "Dominant Political Party by Zip Code: Donation Proportion",
               legend     = "Dem (+); Rep (-)",
               num_colors = 1) + coord_map()
```

### Description Two
On the above map, the darker zip codes code as Democratic and the lighter ones
as Republican. As you can see, the Metro Detroit area and a majority of the coastal
zip codes are Democratic while the Republican patches seem to be spread out around
the state and tend to be landlocked.


## Plot Three

### Support for Hillary vs. Trump over Time (segmented by gender)

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

women <- subset(contrib, gender == 'female' & cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.')) %>% 
  group_by(contrib_month_year, community_type, cand_nm) %>% 
  summarise(total = n())

p1 <- ggplot(data=subset(women,!is.na(community_type)), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~community_type,scales="free",ncol = 1) +
  theme(legend.position="bottom",legend.text=element_text(size=8)) +
  ggtitle("Female")

men <- subset(contrib, gender == 'male' & cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.')) %>% 
  group_by(contrib_month_year, community_type, cand_nm) %>% 
  summarise(total = n())
  

p2 <- ggplot(data=subset(men,!is.na(community_type)), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~community_type,scales="free",ncol = 1) + 
  theme(legend.position="bottom",legend.text=element_text(size=8)) +
  ggtitle("Male")


grid.arrange(p1,p2,ncol=2, top=textGrob("Hillary vs. Trump Support by Date", gp=gpar(fontsize=15,font=8)))
```

### Description Three
Finally, I investigate how support for Hillary vs. Trump differed by gender and
community type as a function of time. In general female voters favored Hillary,
while male voters favored Trump. In terms of community type, city dwellers were
more likely to support Hillary while rural Michiganders supported Trump. Rural
women and urban men are particularly interesting since they combinations of
opposing demographic leanings. They also have extremely similar contribution
profiles. In general they leaned toward supporting Hillary, but had a moment
in July (during convention season and the DNC email scandal) when they were
more supportive of Trump.

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.
