2016 Presidential Campaign Contributions in Michigan 
========================================================

## by Liz Kalina

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(gender)
library(tidyr)
library(devtools)
library(ade4)
library(data.table)
library(grid)
library(choroplethrZip)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Contribution Data for Michigan
header <- read.table("2016-MI.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
contrib <- read.table("2016-MI.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(contrib) <- header

# Load Candidate Info (i.e. Party Affiliation)
header <- read.table("candidate-list.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
cand_info <- read.table("candidate-list.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(cand_info) <- header

# Load Unemployment Dataset
header <- read.table("unemployment-zipcode.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
unemployment <- read.table("unemployment-zipcode.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(unemployment) <- header

# Load Population Density Dataset
header <- read.table("population-density-zipcode.csv", nrow = 1, stringsAsFactors = FALSE, sep = ",")
pop_density <- read.table("population-density-zipcode.csv", skip = 1, stringsAsFactors = FALSE, sep = ",")
names(pop_density) <- header

# Remove NA column(s)
contrib <- contrib[colSums(!is.na(contrib)) > 0]
```

This report explores a dataset containing campaign contributions to all
candidates in the 2016 presidential election by Michigan residents. As a former 
Michigander, I'm particularly interested in Michigan's current political climate.

## Univariate Plots Section

#### Dataset Characteristics

```{r echo=FALSE, message=FALSE, warning=FALSE, Dataset_Characteristics}
str(contrib)
```

#### Unique Candidate Count

```{r echo=FALSE, message=FALSE, warning=FALSE, Unique_Candidates_Count}
length(unique(contrib$cand_nm))
```

Our dataset consists of 18 variables, with 163,765 observations and 24 individual 
candidates that residents contributed to.

#### Summary of Contributions to Each Candidates

```{r echo=FALSE, message=FALSE, warning=FALSE, Contributions_by_Candidate}
cands <- group_by(contrib, cand_nm)
cand_summary <- summarize(cands, total = length(contb_receipt_amt),
                          avg = mean(contb_receipt_amt, na.rm = TRUE), 
                          median = median(contb_receipt_amt), 
                          max = max(contb_receipt_amt))

ordered_cands <- cand_summary[order(cand_summary$total, decreasing=TRUE) , ]
ordered_cands
```

The top Democratic candidates (Hillary Clinton, Bernie Sanders) received the 
largest number of donations, while the top three Republican candidates (Donald 
Trump, Ted Cruz, Ben Carson) received fewer overall donations. The median 
donation per Republican candidate was roughly double that of the Democrats.

#### Summary of Contributions by City

```{r echo=FALSE, message=FALSE, warning=FALSE, Contributions_by_City}
city <- group_by(contrib, contbr_city)
city_summary <- summarize(city, total = length(contb_receipt_amt),
                          avg = mean(contb_receipt_amt, na.rm = TRUE), 
                          median = median(contb_receipt_amt), 
                          max = max(contb_receipt_amt))

city_summary[order(city_summary$total, decreasing=TRUE) , ]
```
The five cities with the largest number of donations were Ann Arbor, Grand 
Rapids, Detroit, Kalamazoo and Lansing. Some cities, such as Bloomfield Hills
and Grosse Pointe, had much larger median donations, which aligns with my 
understanding of the wealth in these communities.

### Contribution Histograms

#### Overall Distribution

```{r echo=FALSE, message=FALSE, warning=FALSE, Overall_Histogram}
ggplot(aes(x = contb_receipt_amt), data = contrib) + 
  geom_histogram(binwidth=5) +
  scale_x_sqrt(limits = c(0,2800))
```

A vast majority of donations were less than $500, which shows that it's not 
just rich donors making a majority of the contributions at least in the context 
of individual donations.

#### Contributions less than $100

```{r echo=FALSE, message=FALSE, warning=FALSE, Less_100}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt < 100)) + 
  geom_histogram(binwidth=10) +
  scale_x_continuous(limits = c(0, 100))
```

For donations less than $100, you can notice peaks around $25, $50 and $75
which aligns with suggested donation amounts incremented by $25.

#### Contributions less than $400

```{r echo=FALSE, message=FALSE, warning=FALSE, Less_400}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt < 400)) + 
  geom_histogram(binwidth=50) +
  scale_x_continuous(limits = c(0, 400))
```

#### Contributions greater than $1000

```{r echo=FALSE,message=FALSE, warning=FALSE,  Greater_1000}
ggplot(aes(x = contb_receipt_amt), data = subset(contrib,contb_receipt_amt > 1000)) + 
  geom_histogram(binwidth=750) +
  scale_x_continuous(limits = c(1000, 6000)) +
  scale_y_continuous(limits = c(0, 1500))
```

According to the Federal Election Commission, individual contribution limits are:

> $2,700 per election to a Federal candidate or the candidate's campaign committee. Notice that the limit applies separately to each election. Primaries, runoffs and general elections are considered separate elections.

Since presidential campaigns have primaries and general elections, you can see 
this reflected in the clustering of contribution counts at $2700 and $5400. 

### Candidates with the Most and Fewest Donations

```{r echo=FALSE, message=FALSE, warning=FALSE, Most_Fewest_Donations}
p1 <- ggplot(aes(y=total, x= cand_nm),data=head(ordered_cands, 5)) + 
        geom_bar(stat="identity") + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle("Most Contributions")

p2 <- ggplot(aes(y=total, x= cand_nm),data=tail(ordered_cands, 5)) + 
        geom_bar(stat="identity") + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle("Least Contributions")


grid.arrange(p1,p2,ncol=2)
```

Above is a scale comparison of the candidates with the most and least 
contributions. The total contribution difference is vast; top candidates can
have more than half a million contributions, while four of the five smallest
candidates have less than 10 donations.

### Contributor Occupations and Employers

#### By Employer

```{r echo=FALSE, message=FALSE, warning=FALSE, Employmer_Stats}
contrib %>% 
  group_by(contbr_employer) %>% 
  summarise(total = n()) %>%
  arrange(desc(total))
```

#### By Occupation

```{r echo=FALSE, message=FALSE, warning=FALSE, Occupation_Stats}
contrib %>% 
  group_by(contbr_occupation) %>% 
  summarise(total = n()) %>%
  arrange(desc(total))
```

Employment Observations

+ Retired people are the most frequent contibutors by far
+ Employees of universities, automakers and hosptial systems are frequent contributors
+ There are a large number of contributions without employer and occupation info


### Contribution Timing

```{r echo=FALSE, message=FALSE, warning=FALSE, Format_Date}
## Add Date to Contributions
contrib$contrib_date <- as.Date(contrib$contb_receipt_dt, '%d-%b-%y')
contrib$contrib_month_year <- format(contrib$contrib_date, '%Y-%m')
```

```{r echo=FALSE, message=FALSE, warning=FALSE,  Count_by_Date}
## Contribution Counts by Date
contrib_by_month_year <- contrib %>% 
  group_by(contrib_month_year) %>% 
  summarise(total = n())

ggplot(data=contrib_by_month_year, aes(x=contrib_month_year, y=total, group=1)) +
  geom_line() +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```

From the plot above, you can see a gradual climb in donations during 2015 and a 
steep climb from January to March 2016. There are three contribution peaks in 
2016 — in March, July and October — each coinciding with an election event. There 
was the Michigan Primary on March 8th, the Democratic and Republican national 
conventions at the end of July and the general election on November 8th.


## Univariate Summary

To me, the most interesting features in this dataset are 'cand_nm', 'contbr_nm', 
'contbr_zip', 'contb_receipt_amt' and 'contb_receipt_dt.' I can use 
'contb_receipt_amt' and 'contb_receipt_dt' for each contribution to investigate
temporal and candidate popularity trends. With 'contbr_nm' and 'contbr_zip', I can
infer information about the contributor such as geographic location and gender.

A few takeaways from my analysis so far are that Democratic candidates received
more contributions than Republicans and that a majority of contributions are
small (less than $500). Other interesting things I noticed were that contribution
amounts cluster around campaign maxs for larger donors and contributions peak 
around key election events (like primaries, conventions or scandals).

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Add_Pol_Party}
## Add Political Party Affiliation
contrib <- merge(contrib, subset(cand_info[c("cand_id", "cand_pty_aff")]), by = "cand_id")
## This political party data comes from the Federal Election Commission website.
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Split_Name}
## Split Full Name Column into First and Last Name
contrib <- separate(contrib, contbr_nm, c("last_name", "first_name"), sep = ",", extra="merge", fill="right")
contrib <- data.frame(lapply(contrib, trimws))
contrib$clean_first_name <- gsub("([A-Za-z]+).*", "\\1", contrib$first_name)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Add_Gender}
## Add Gender to Contributors
gender_lookup <- gender(c(unique(contrib$clean_first_name)))
contrib <- left_join(contrib,gender_lookup,by = c("clean_first_name"="name"))

contrib_by_date_w_gender <- subset(contrib,proportion_male > 0.75 | proportion_female > 0.75) %>% 
  group_by(contrib_month_year, gender ) %>% 
  summarise(total = n())
```

#### Add Gender to Contributors

In order to determine each contributors gender, I used the 'gender' package in R,
which encodes gender based on names and dates of birth using historical datasets.
I decided to use a threshold of 75% certainty to assign a gender to a contributor.

### Political Party Investigation

#### Contributor Gender vs. Candidate Political Party
```{r echo=FALSE, message=FALSE, warning=FALSE, Gender_v_Party}
ggplot(aes(x=cand_pty_aff),
       data=subset(contrib,!is.na(gender) & cand_pty_aff %in% c('Republican Party','Democratic Party'))) +
  geom_histogram(stat='count',aes(fill=cand_pty_aff)) +
  facet_wrap(~gender)
```

Both men and women in Michigan donate more to Democrats than Republicans;
however, women are significantly more likely to donate Democratic than men.

#### Contributions Counts by Date and Political Party
```{r echo=FALSE, message=FALSE, warning=FALSE, Date_v_Party}
contrib_by_date_w_party <- contrib %>% 
  group_by(contrib_month_year, cand_pty_aff ) %>% 
  summarise(total = n())

ggplot(data=na.omit(subset(contrib_by_date_w_party,cand_pty_aff != " ")), aes(x=contrib_month_year, y=total, group=cand_pty_aff)) +
  geom_line(aes(color = cand_pty_aff)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```

While Republicans actually donated more prior to 2016, the Democratic candidates
received the overall majority of contributions. There were very few donations
that were Libertarian or not marked for a particular party.

#### Republican Candidate Contribution Counts by Date
```{r echo=FALSE, message=FALSE, warning=FALSE, Rep_Counts}
reps <- subset(contrib, cand_pty_aff == 'Republican Party') %>% 
  group_by(contrib_month_year, cand_nm) %>% 
  summarise(total = n())

ggplot(data=subset(reps,total > 10), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)])
```

I think this plot is really interesting because it shows that Trump had almost 
no financial contributions prior to the Republican Primary on March 8, 2016. The 
leaders before the primary were Ben Carson and Ted Cruz. Trump's support 
peaked at the Republican National Convention at the end of July and then
steadily declined until the month before the general election when it had a
small resurgence.

## Bivariate Summary

In this section, I added political affiliation to candidates and gender to
contributors, which allowed me to see the relationship between these features
and campaign contributions. There is a very strong correlation between female 
gender and Democratic Party contributions. Other discoveries include how 
cluttered the Republican race was prior to primaries, the lack of Trump support 
prior to the Republican Primary and the strong Bernie support before the 
Democratic National Convention.

# Multivariate Plots Section

#### Hillary vs. Trump over Time

```{r echo=FALSE, message=FALSE, warning=FALSE, Hill_v_Trump}
hill_v_trump <- subset(contrib, cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.') & !is.na(gender)) %>% 
  group_by(contrib_month_year, gender, cand_nm) %>% 
  summarise(total = n())

ggplot(data=hill_v_trump, aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_bar(aes(fill = cand_nm),position = "dodge",stat="identity") +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~gender)
```

Donation trends to Hillary vs. Trump were uniform in shape, but not in scale
across the genders and time. One key difference is that women always donated 
more to Hillary than Trump, while men donated more to different candidates at 
different points in time. In the months right before the general election, men
actually made more donations to Hillary!

### Contribution Amount Distribution

#### Top 10 Candidates

```{r echo=FALSE, message=FALSE, warning=FALSE, Top_10}
top_10 <- head(ordered_cands, 10)$cand_nm

ggplot(aes(x = cand_nm,y=contb_receipt_amt), 
       data=subset(contrib,cand_nm %in% top_10 & contb_receipt_amt > 0 & contb_receipt_amt <= 2700)) + 
  geom_boxplot(aes(fill=cand_pty_aff)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Contribution Observations

+ The Republicans receive larger contributions than Democrats
+ Hillary has a low interquartile range, but still has many large contributions 
+ Jeb Bush has comically large contributions 

#### By Political Party and Gender

```{r echo=FALSE, message=FALSE, warning=FALSE, Party_v_Gender}
contrib$contb_receipt_amt <- as.numeric(as.character(contrib$contb_receipt_amt))

ggplot(aes(x = cand_pty_aff,y=contb_receipt_amt), data=subset(contrib,!is.na(gender) & cand_pty_aff %in% c('Republican Party','Democratic Party') & contb_receipt_amt > 0 & contb_receipt_amt <= 2700)) + 
      geom_boxplot(aes(fill=cand_pty_aff)) +
      facet_wrap(~gender)
```

From this boxplot, you can see that the interquartile range for contribution 
amounts is less than $100 across gender/political party and that Republican men 
give the most per contribution.

### Determine Population Density and Unemployment Rate of Contributor Zip Codes

```{r echo=FALSE, message=FALSE, warning=FALSE, Add_Unemp_Pop_Den}
# Join the unemployment and population density datasets to contributor table
contrib$contbr_zip <- as.integer(substr(contrib$contbr_zip, 0, 5))
contrib <- left_join(contrib,unemployment,by = c("contbr_zip"="Zip"))
contrib <- left_join(contrib,pop_density,by = c("contbr_zip"="Zip/ZCTA"))

# Convert unemployment string (i.e. '8%') to numeric equivalent
contrib$`Unemp. Rate` <- as.numeric(sub("%", "", contrib$`Unemp. Rate`))
```

These two datasets were available on the Splitwise blog as downloads. According
to the author, this data came from:

> Two different Census APIs (the Decennial Census 2010 and the ACS 5-year 2007-2011), 
combined with the square-footage by ZCTA listings from the 2013 U.S. Gazetteer Files.

I believe that understanding the community type and economic prospects will be 
a strong indicator of candidate preference.

### Classify Contributor Community Type based on Population Density 

```{r echo=FALSE, message=FALSE, warning=FALSE, Assign_Community_Type}
contrib$community_type <- cut(contrib$`Density Per Sq Mile`, breaks = c(0, 102, 2213,10000), labels = c('rural', 'suburban', 'urban'), right=FALSE)
```

Classification (People per Square Mile):

+ rural: < 102
+ suburban: 103 to 2,213
+ urban: > 2213

These cutoffs are based on an analysis presented on the FiveThirtyEight blog:

> Our analysis showed that the single best predictor of whether someone said his or her area was urban, suburban or rural was ZIP code density. Residents of ZIP codes with more than 2,213 people per square mile typically described their area as urban. Residents of neighborhoods with 102 to 2,213 people per square mile typically called their area suburban. In ZIP codes with fewer than 102 people per square mile, residents typically said they lived in a rural area.


#### Party Affiliation vs. Unemployment vs. Population Density

```{r echo=FALSE,message=FALSE, warning=FALSE,Multivariate_Plot}
ggplot(aes(y=`Unemp. Rate`,x=`Density Per Sq Mile`), data=contrib) +
    geom_point(aes(color=cand_pty_aff),alpha = 1/100, size=3) +
      scale_color_brewer(type = 'div',
      guide = guide_legend(title = 'Party Affiliation', reverse = T,
      override.aes = list(alpha = 1, size = 2))) +
      scale_x_sqrt() +
      xlim(0,8000) +
      ylim(0,35)
```

In general, Democratic contributors appear to live in areas with higher population
density and higher unemployment while Republicans seem to live in areas with 
average unemployment and lower to moderate population density.

#### Correlation: Population Density, Gender, Party Affiliation

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation_One}
feats = c('gender','cand_pty_aff')
for (f in feats){
  df_dummy = acm.disjonctif(contrib[f])
  contrib = cbind(contrib, df_dummy)
}

cor_1 <- cor(na.omit(contrib[,c('Density Per Sq Mile','gender.male','gender.female','cand_pty_aff.Democratic Party','cand_pty_aff.Republican Party')]))

upper <- round(cor_1,2)
upper[upper.tri(cor_1)] <- ''
upper <- as.data.frame(upper)

upper
```

Correlation Observations

+ There are minor correlations between:
  + Democrats: higher population density, female
  + Republicans: lower population density, male


```{r echo=FALSE, message=FALSE, warning=FALSE, Group_Party_Zip}
## Group by zip code
group_by_zip <- contrib %>% 
  group_by(contbr_zip) %>% 
  summarise(total=n(),value=median(contb_receipt_amt))

setnames(group_by_zip, "contbr_zip", "region")
group_by_zip$region <- as.character(group_by_zip$region)

## Group by political party
group_by_party <- contrib %>%
 group_by(contbr_zip,cand_pty_aff) %>% 
 summarise(n= n()) %>%
 ungroup() %>%
 spread(cand_pty_aff, n, fill=0) %>%
 mutate(total = `Republican Party` + `Democratic Party`,
        rep_percent = (`Republican Party`/total),
        dem_percent = (`Democratic Party`/total)) %>%
 mutate(value = ifelse(`Republican Party`>`Democratic Party`, -(rep_percent),
               ifelse(`Republican Party`<`Democratic Party`, (dem_percent), NA)))

setnames(group_by_party, "contbr_zip", "region")
group_by_party$region <- as.character(group_by_party$region)
```

#### Donation Amounts vs. Zip Code

```{r echo=FALSE, message=FALSE, warning=FALSE, Donation_Map}
don_amt <- subset(group_by_zip,total > 10 & !is.na(region), c('region','value'))

zip_choropleth(don_amt, 
               state_zoom = "michigan", 
               title      = "Median Donation by Zip Code",
               legend     = "Median Donation",
               num_colors = 4) + coord_map()
```

The geography of median donation seems to be relatively evenly distributed across
the state with the exeception of Metro Detroit which is almost uniformly less 
than $30. The black regions are excluded zip codes since they have less than 
10 donations.

#### Correlation: Median Donation, Total Donations, Party Leaning

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation_Two}
amt_pty_cor <- full_join(group_by_zip,group_by_party,by = 'region')

setnames(amt_pty_cor, "value.x", "median donation")
setnames(amt_pty_cor, "total.x", "total donations")

pairs(na.omit(amt_pty_cor[,c('median donation','total donations','dem_percent','rep_percent')]))
```

The above scatterplot matrix shows each zip code as a point. This plot simply
reinforces that Democrats live in more densely populated areas than Republicans,
since the Democratic zip code have more total donations than the Republican zips.

## Multivariate Summary

In this section, I investigated support for Hillary vs. Trump over time, the
size of campaign contributions for Democrats vs. Republicans as well as the 
relationship between population density and unemployment to party affiliation. I 
observed that Democratic indicators include high population density and female 
gender, while Republican indicators are low population density and male gender. 
Other observations include that Republicans (particularly Republican men) give 
larger donations and although men favored Trump to Hillary, men actually made 
more contributions to Hillary in the last few months leading up to the general 
election.

------

# Final Plots and Summary

## Plot One

### Hillary vs. Bernie

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
hill_v_bern <- subset(contrib, cand_nm %in% c('Clinton, Hillary Rodham','Sanders, Bernard')) %>% 
  group_by(contrib_month_year, cand_nm) %>% 
  summarise(total = n())

ggplot(data=hill_v_bern, aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_bar(aes(fill = cand_nm),position = "dodge",stat="identity") +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,3)]) +
  xlab("Contribution Date") +
  ylab("Contribution Count") +
  ggtitle("Hillary vs. Bernie Support by Date")
```

### Description One

The above plot shows that Bernie Sanders actually had more support (based on 
number of contributions) from 2015 to well after the Democratic Primary. It was 
not until June 2016 that Hillary surpassed Bernie in number of donations, and not
until after the DNC that Bernie support fully dried up. This coincides with my 
understanding of Bernie as a grassroots candidate with widespread and loyal 
supporters.

## Plot Two

### Party Leanings based on Zip Code


```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
pty_ct <- subset(group_by_party,total > 5 & !is.na(region), c('region','value'))

zip_choropleth(pty_ct, 
               state_zoom = "michigan", 
               title      = "Dominant Political Party by Zip Code: Donation Proportion",
               legend     = "Dem (+); Rep (-)",
               num_colors = 1) + coord_map()
```

### Description Two

On the above map, the darker zip codes code as Democratic and the lighter ones
as Republican. As you can see, the Metro Detroit area and a majority of the coastal
zip codes are Democratic, while the Republican patches are spread out around
the state and tend to be landlocked.

For the politcal party calculations, I select the dominant party (based on number
of contributions) for each zip code and calculate the percentage of total 
contributions that party receives.

## Plot Three

### Support for Hillary vs. Trump over Time (segmented by gender)

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
women <- subset(contrib, gender == 'female' & cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.')) %>% 
  group_by(contrib_month_year, community_type, cand_nm) %>% 
  summarise(total = n())

p1 <- ggplot(data=subset(women,!is.na(community_type)), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~community_type,scales="free",ncol = 1) +
  theme(legend.position="bottom",legend.text=element_text(size=8)) +
  xlab("Contribution Date") +
  ylab("Contribution Count") +
  ggtitle("Female")

men <- subset(contrib, gender == 'male' & cand_nm %in% c('Clinton, Hillary Rodham','Trump, Donald J.')) %>% 
  group_by(contrib_month_year, community_type, cand_nm) %>% 
  summarise(total = n())
  

p2 <- ggplot(data=subset(men,!is.na(community_type)), aes(x=contrib_month_year, y=total, group=cand_nm)) +
  geom_line(aes(color = cand_nm)) +
  scale_x_discrete(breaks = unique(contrib_by_month_year$contrib_month_year)[seq(1,28,6)]) +
  facet_wrap(~community_type,scales="free",ncol = 1) + 
  theme(legend.position="bottom",legend.text=element_text(size=8)) +
  xlab("Contribution Date") +
  ylab("") +
  ggtitle("Male")


grid.arrange(p1,p2,ncol=2, top=textGrob("Hillary vs. Trump Support by Date/Gender/Community Type", gp=gpar(fontsize=13,font=8)))
```

### Description Three

Finally, I investigate how support for Hillary vs. Trump differed by gender and
community type as a function of time. In general female voters favored Hillary,
while male voters favored Trump. In terms of community type, city dwellers were
more likely to support Hillary while rural Michiganders supported Trump. 

Rural women and urban men are particularly intriguing since they're combinations 
of opposing political leanings demographics. They also have extremely similar 
contribution profiles. In general, they leaned toward supporting Hillary, but 
both had a moment in July (during convention season and the DNC email scandal) 
when they were more supportive of Trump.

------

# Reflection

### Features Examined

contribution

+ amount
+ date

candidate

+ political party

contributor

+ employer
+ occupation
+ gender

zip code

+ population density
+ community type
+ unemployment rate
+ political party leaning

### Takeaways

I learned so much through this exploration! One discovery that came up 
throughout my analysis was that Michigan residents donated more (based on number
of contributions) to Democratic candidates, but made larger individual 
contributions to Republicans. Another discovery was that Democrats tend to be 
urban, female and located on the coasts, while Republicans tend to be rural, 
male and landlocked. A final takeway was how clearly you can see campaign events 
(like primaries, conventions and scandals) in the data.

#### Surprises

After the 2016 election, I heard so much about how women voted for Trump in 
droves, which supposedly won the election for him. What surprised me from this 
analysis is that although this may be true from a voter turnout level, it's not 
what is reflected in the individual contributor data. 

Other things that suprised me surrounded the campaign timeline and candidate 
support. Prior to this investigation, I didn't realize what strong support 
Bernie enjoyed between the primaries and DNC, or what minimal support Trump had 
in comparison to other Republican candidates before the Republican Primary. 
Another timeframe surprise was that Michigan men actually made more total 
contributions to Hillary than Trump in the months leading up to the general 
election.

#### Struggles

The main struggle that I encountered during this project was learning the 
idiosyncrasies of R and ggplot. I also encountered several situations where 
numeric columns were encoded as factor data, which led to issues where I 
couldn't plot correctly because the values weren't recognized as continuous. I 
solved these factor problems by simply converting to numeric values.

#### Next Steps

As next steps, I would like to more closely examine how community type (rural, 
suburban, urban) affects contribution outcomes using race, income and voting 
statistics. Once these features are included, I would like to delve into 
individual cities and regions within Michigan to better understand campaign 
donation and voting tendencies.



